---
title: "SAGE Gait Study - Sub-Saharan Africa"
author: '*Boadiwaa*'
date: "Created on `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE,warning = FALSE)
library(knitr)
library(haven)
library(labelled)
library(foreign)
library(gtsummary)
library(ggsci)
library(flextable)
library(qpcR)
library(ggpubr)
library(ggbeeswarm)
library(ggthemes)
library(ggrepel)
library(gt)
library(tidyverse)

knitr::read_chunk('edits_gait.R')

```

### **RESULTS**

### **Data Analysis**

This report is based on the analysis of  the individual questionnaires for  Ghana and South Africa. 

#### Data Cleaning




### Overview of Demographics and Socioeconomic Indices for Sub Saharan Africa (Ghana and South Africa combined)

The majority of respondents (~84% representing 8564 respondents) were aged 50 years and above, and 1224 were from the control group (ages 18 to 49).  This represents ~ 1:7 ratio of control group to study group.

**Empirical results from study group:**

Observed median age of  62   and 61 were recorded for Ghana and South Africa respectively. The age group with the lowest count was for 80 years and above having 9.8% and 6.5% of the respective totals. There was equal gender representation in Ghana entries but more females (57%) from South Africa.  The majority (59%) of respondents in Ghana resided in rural areas whereas for South Africa, the opposite (33%) was noted. The majority of respondents were either cohabiting or “currently married” in both countries, with the next highest proportion widowed. The proportion of respondents who were separated/divorced in in Ghana was about twice the proportion in South Africa, and about nine times the proportion of respondents who had never married in Ghana had never married in South Africa. Income was almost equally distributed across the 5 quintiles.

```{r, a}

```

### Overview of Health-related Indices for Sub-Saharan Africa (Ghana and South Africa combined)  

The majority of respondents had no history of any of the comorbidities studied, except in South Africa where majority of respondents had a history of “Arthritis and Musculoskeletal Pains”.

```{r, com}

```

### Distribution of Impaired Gait Speed Amongst the Study Group

A Gait Speed of 0.1m/s was used as the cut-off for impaired gait speed, whether normal or rapid, as seen in many past studies.

```{r, aig}

```  


### In-country Differences

#### Ghana Data

Focusing on Ghana, the median age was 64 for females and 61 for males. The majority of males were "currently married" whereas the majority of females were widowed. Males had a slightly higher normal gait speed than females. With rapid gait speed, the gender difference increased to 0.21. The majority of males had a history of tobacco use whereas majority of females had no history.

#### South Africa Data

Focusing on South Africa, the median age was 61 for females and 60 for males. The majority of males were "currently married" whereas the majority of females were widowed. Males had a slightly higher normal gait speed than females. With rapid gait speed, the gender difference increased slightly further. Differences in gait speed were also noted based on ethnicity.Those identifying as "White" had a significantly higher gait speed as compared to those of other ethnicities.This observation was irrespective of income level. Income levels did not affect the median gait speed of respondents. 

```{r, asa}

``` 

```{r, ieg}

``` 

### In-Country and Gender Differences: Socio-Demographic Characteristics

More females in  South Africa had more years of education as compared to males. The opposite was seen in Ghana. In both countries, more men were "currently married/cohabiting" as compared to women, and a higher proportion of females were widowed or "separated/divorced" as compared to males. In South Africa, more women had never married as compared to men. In Ghana, there was no gender difference.

```{r, b}

```

### In-Country and Gender Differences: Co-Morbidities

Generally, more people did not have a history of alcohol use in both countries. For those who did though, slightly more women had a positive history as compared to men. 

For tobacco use, almost four times more women in South Africa had a positive history compared to women in Ghana. there was not much difference between men across the two countries.

In total, women had more positive histories of comorbidities as compared  to men.

```{r c}

```

### Comparison of Gait Speed Across Demographic Variables

The table below shows the multiple regression results with the gait speed as the outcome variable and the various demographics variables as the independent variables.

```{r d}

```


### Comparison of Gait Speed Across Health-Related variables

<!-- Obesity,defined as a Body-Mass Index (BMI) of 30kg/m2 and above, decrease in mean gait speed which was considered statistically significant. Being underweight (BMI <18.5kg/m2) was associated with a reduction in mean gait speed by 0.03 (statistically insignificant).   -->

History of alcohol use does not show a significant difference in gait speed of those who responded yes relative to those who responded no for both countries when all other variables are held constant.

There is a 0.04 increase (significance of 0.004) in gait speed for persons with a positive history of tobacco use relative to those with no history of tobacco use in Ghana. There is an increase of 0.01 in South Africa but not significant.  This is when all other variables are held constant.

History of Depression reduced gait speed significantly for "yes" relative to "no" in Ghana but not in South Africa.

A positive history of Diabetes Mellitus reduces gait speed: no significant reduction in Ghana but significant in South Africa.


```{r, e}

```

#### Association of health indices with gait speed  

Using the 12-item World Health Organization’s Disability Assessment Schedule, version 2 (WHODAS-II), respondent’s functional ability was measured. The 12-item questionnaire measured the activities of daily living(ADLs) and instrumental activities of daily living (IADLs).

The WHODAS-related variables are a series of questions that start with the following preamble: “Overall in the last 30 days, how much difficulty…”, followed by various parameters/tasks.
The respondents' answers were re-coded with **0 equivalent to "none", 1 equivalent to "mild", 2 to "moderate", 3 to "severe" and 4 to "extreme".** Across the 12 variables, a respondent could score a **minimum of 0 and a maximum of 48.** The table below compares the 12-item WHODAS variables in general, and specifically the three questions which deal with the ADLs, as well as height, and BMI and gait speed.

```{r, n}
 
```

### Trend of Gait Speed With Age
It can be seen clearly in the graphs below that there is a decline in gait speed with increasing age:
```{r, g}
 
```

### Zooming In To 75+
The black line represents the linear regression model with the shaded area representing the 95% confidence interval. 
```{r, h}

```

### Respondents' subjective assessment of pain, health and energy levels

### Association of Health Rating with gait speed

The graph below shows respondents' subjective assessment of health level and it's association with gait speed. The line in the box represents the median, and the entire box represents the interquartile range. The density of the surrounding dotted area represents the number of respondents' who chose that answer as their health rating.

```{r, ps}

```

```{r, pe}

``` 