---
title: "SAGE Gait Study - Sub-Saharan Africa"
author: '*Boadiwaa*'
date: "Created on `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE,warning = FALSE)
library(knitr)
library(tidyverse)
library(foreign)
library(ggpubr)
library(gtsummary)
library(ggsci)
library(flextable)
library(grafify)
knitr::read_chunk('edits_gait.R')

```

### Introduction
The Study on global AGEing and adult health (SAGE) is part of a Longitudinal Survey Programme in the World Health Organization (WHO)’s  Multi-Country Studies unit. The main SAGE surveys compiled longitudinal information on the health and well-being of adult populations and the ageing process from nationally representative samples in six countries (China, Ghana, India, Mexico, Russian Federation and South Africa).  Wave 1 of SAGE was conducted between 2007 and 2010. This report focuses on the two countries from Sub-Saharan Africa: Ghana and South Africa. 

### **Objective of report**

This report aims to find associations between gait speed, and various sociodemographic, health and quality of life parameters.

### **Method**
The information in this section, with the exception of the Data Analysis subsection, is entirely derived from the SAGE WAVE 1 Ghana and SAGE WAVE 1 South Africa national reports.

#### **Sample**

#### Ghana
Primary sampling units (PSUs) were selected based on proportional allocation by size and then stratified by region and location (urban/rural). Each enumeration area (EA) was selected independently
within each stratum. 250 EAs were selected based on the availability of respondents aged 50 years and above. Within each EA, 20 households with one or more 50+ individuals and four households with members aged 18–49 years were randomly selected. Field work and data entry were undertaken between May 2007 and June 2008.

#### South Africa
A two-stage probabilistic sample design was used. The first stage of sampling was the selection of primary
sampling units (PSUs) which is a probabilistic sample of 1000 enumeration areas (EA) drawn from the South African National Census. For the SAGE study in South Africa, a total sample of 600 EAs was drawn from the master sample and used as the PSUs. The EAs were then selected with a probability proportional to size, with the estimated number of people aged 50 years or older in each EA as a measure of size. Thus, EAs with a larger number of people aged 50 years or older had a higher chance of being selected.

#### **Data Collection Protocol**

#### Ghana
Six types of questionnaires were used in the SAGE study:

- a household questionnaire
- an individual questionnaire
- a proxy respondent questionnaire
- a verbal autopsy questionnaire
- a proxy validation questionnaire
- a retest questionnaire

Data collection was done manually during face-to-face interviews in respondents’ homes. A team consisted of four interviewers and one supervisor assigned to one PSU. Each interviewer was to complete interviews with two respondents per day. Completed questionnaires were organized and documented centrally before data entry. Two centres were organized in Ghana for data entry and then the final product was sent to the WHO.

#### South Africa
Four types of questionnaires were used in the SAGE study:

- a household questionnaire
- an individual questionnaire
- a proxy respondent questionnaire
- a verbal autopsy questionnaire

Data collection was done in a CSPro software application provided by WHO, during face-to-face interviews in respondents’ homes. A team consisted of three to five interviewers and one supervisor. On average each interviewer completed two households and two individual interviews per day. The final raw data was sent to the WHO for further cleaning and weighting.

#### **Survey and Data Quality**

**Response rates** are a common indicator of survey quality. 

For the Ghana survey:  

The response rate for the individual questionnaire ranged between 92.1% for women and 97.0% for men.

For the South Africa survey:  

The response rate was only calculated for those aged 50 years or older and was 60%.

**Age heaping** is considered to be a measure of data quality and consistency. Myers' blended index is a measure of age heaping and digit preference. If Myers’ blended index is over 60, age heaping is very severe and the data quality is poor.

For the Ghana survey:  

Myers’ blended index was 11.9, which indicates that a minimum of 11.9% of the population reported ages with an incorrect final digit. The index value is very low with some evidence of heaping on end digits 0 and 5.

For the South Africa survey:  

The Myers'blended index was 3.3, with no evidence of age heaping.

From the above metrics, it can be concluded that the data to be analyzed is of good quality.


### **Data Analysis**

Only data entered from the individual questionnaires from Ghana and South Africa were analyzed for this report.

#### Data Cleaning

5573 entries were collected from Ghana and 4227 entries from South Africa. The outcomes of interest for this study were the normal and rapid gait speeds. For comparisons between independent variables, the normal gait speed was used as the dependent variable.

After the demographics table, only entries with time above 0 seconds(s) for both normal and rapid gait were selected, reducing the number of entries to 4835 for Ghana and 3666 for South Africa. There was a huge gap between the third quartile and the maximum for both gait speeds which strongly suggested data collection errors. Upon further checks, a range of 1.0 to 12.0s was used for the normal gait speed for both Ghana and South Africa. 1.2 to 13.0s and 1.0 to 13.0s was used as the range for the rapid gait speed for Ghana and South Africa respectively and thus, leaving 136 outliers from Ghana and 119 from South Africa, and  86 outliers from Ghana and 58 outliers from South Africa for the normal and rapid gait speeds respectively. Other minor data cleaning and validation processes such as selection of relevant variables, and finding the average blood pressure and pulse reading from a set of three were carried.

The two data sets were then combined yielding a total of 8501 entries.


### Overview of Demographics, Socioeconomic and Health Indices for Sub Saharan Africa (Ghana and South Africa combined)

As shown in the table below, the majority of respondents (~84%) were aged 50 years and above, and 51% were at least 60 years old..Median age was 60 for both Ghana and South Africa. The lowest count for the age groupings was for 80 years and above. The majority of respondents from Ghana were male, and from South Africa: female. The majority of respondents did not provide an answer when asked the number of years of (formal) education they had received. For those who did, majority of those from Ghana had completed a minimum of 10 and a maximum of 12 years of education and the majority for South Africa had a minimum of 7 and a maximum of 9 years.

The majority of respondents in Ghana resided in rural areas whereas for South Africa, the opposite was noted. The majority of respondents were "currently married" in both countries, with the next highest proportion widowed. About twice the proportion of respondents who were separated/divorced in South Africa were separated/divorced in Ghana, and about thrice the proportion of respondents who had never married in Ghana had never married in South Africa. The majority of respondents had low income satisfaction.

The majority of respondents in Ghana were of normal weight whereas the majority from South Africa were obese.The majority of respondents had no history of chronic disease or alcohol use. For tobacco use however, the majority from Ghana had a positive history whereas the majority from South Africa had a negative history.





```{r, a}

```

Preliminary analysis of the gait speeds showed that the data distribution is skewed to the right, and thus the median was used as the measure of central tendency. The normal gait speed was the same in both countries, and South Africa had a slightly higher rapid gait speed. the table below shows the median gait speed for both countries and the graph shows the distribution.

```{r, gs}

```

### In-country Differences

#### Ghana Data

Focusing on Ghana, the median age was 60 for females and 58 for males. The majority of males were "currently married" whereas the majority of females were widowed. Males had a slightly higher normal gait speed than females (0.09 difference). With rapid gait speed, the gender difference increased to 0.2. The majority of males had a history of tobacco use whereas majority of females had no history.


```{r, b}

```

#### South Africa Data

Focusing on South Africa, the median age was 60 for females and 59 for males. The majority of males were "currently married" whereas the majority of females were widowed. Males had a slightly higher normal gait speed than females (0.10 difference). With rapid gait speed, the gender difference increased to 0.18.

```{r c}

```

### Comparison of Median Gait Speed Across Demographic Variables

The table below shows the median gait speed with the end-points of the interquartile range in brackets. there is a statistically significant decrease in normal gait speed with increasing age, with a 0.21 decrease in ages 80 and above compared to ages 49 and below. There is no apparent association between years of education and gait speed. Respondents in rural areas had a slightly higher gait speed than those in urban areas (0.02) difference, which was not statistically significant. Having a partner ("currently married"or "cohabiting") was associated with a slightly higher gait speed though this association was not statistically significant. Having low income satisfaction was associated with lower gait speed though this association was not statistically significant.  


```{r d}

```


### Comparison of Median Gait Speed Across Health-Related variables

Obesity,defined as a Body-Mass Index (BMI) of 30kg/m2 and above, was associated with a 0.05 decrease in mean gait speed which was considered statistically significant. Being underweight (BMI <18.5kg/m2) was associated with a reduction in mean gait speed by 0.03 (statistically insignificant).  

Generally, having no history of chronic disease was associated with higher gait speed, with the exception of Diabetes Mellitus, Injury from Road Traffic Accidents and Cataracts. A history of alcohol use was associated with a statistically significant lower gait speed. On the other hand, a history of tobacco use was associated with an insignificant increase in gait speed.


```{r, e}

```

#### Combined Linear Regressions Table
```{r, f}

```

### Association of WHODAS variables and gait speed 

Using the 12-item World Health Organization's Disability Assessment Schedule, version 2 (WHODAS-II), respondent’s functional ability was measured. The 12-item questionnaire measured the activities of daily living (ADLs) and instrumental activities of daily living (IADLs). 

The WHODAS-related variables are a series of questions that start with the following preamble:
"Overall in the last 30 days, how much difficulty…", followed by various parameters/tasks.

The table below gives an overview of the WHODAS questions and answers.

##### **"Overall in the last 30 days, how much difficulty: "**
```{r, m}

```


#### Association of WHODAS score with gait speed

The respondents' answers were re-coded with **0 equivalent to "none", 1 equivalent to "mild", 2 to "moderate", 3 to "severe" and 4 to "extreme".** Across the 12 variables, a respondent could score a **minimum of 0 and a maximum of 48.** As shown in the table below, the highest score was 10 and there is no apparent effect of the score on the gait speed.

```{r, n}

```

Here is a visualization of the WHODAS score in relation to gait speed in those aged 50 years and above:

```{r, wg}
 
```

### Trend of Gait Speed With Age
It can be seen clearly in the graphs below that there is a decline in gait speed with increasing age:
```{r, g}
 
```

### Zooming In To 60+
The black line represents the linear regression model with the shaded area representing the 95% confidence interval. 
```{r, h}

```


### Gender vs Normal and Rapid Gait Speed
The results below suggest that the male gender is associated with slightly higher gait speed:
```{r, i}

```


### Respondents' subjective assessment of pain, health and energy levels
```{r, k}

```

### Association of Pain, Health Rating and Energy Levels with gait speed

There is a statistically significant decrease in gait speed with a poor ranking of respondents'pain levels, health rating and energy levels.

```{r, km}

```

#### Related Graphs
The figures below are graphical representations of the above parameters.
```{r, ps}

```

```{r, l}

```


```{r, o}

```


```{r, p}

```

```{r, q}

```