---
title: "SAGE Gait Study - Sub-Saharan Africa"
author: '*Boadiwaa*'
date: "Created on `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE,warning = FALSE)
library(knitr)
library(tidyverse)
library(foreign)
library(ggpubr)
library(gtsummary)
library(ggsci)
library(flextable)
library(grafify)
library(qpcR)
library(ggstatsplot)
library(ggbeeswarm)
library(ggthemes)
library(ggrepel)

knitr::read_chunk('edits_gait.R')

```

### **RESULTS**

### **Data Analysis**

Only data entered from the individual questionnaires from Ghana and South Africa were analyzed for this report.

#### Data Cleaning

5573 entries were collected from Ghana and 4227 entries from South Africa. The outcomes of interest for this study were the normal and rapid gait speeds. For comparisons between independent variables, the normal gait speed was used as the dependent variable.

After the demographics table, only entries with time above 0 seconds(s) for both normal and rapid gait were selected, reducing the number of entries to 4835 for Ghana and 3666 for South Africa. There was a huge gap between the third quartile and the maximum for both gait speeds which strongly suggested data collection errors. Upon further checks, a range of 1.0 to 12.0s was used for the normal gait speed for both Ghana and South Africa. 1.2 to 13.0s and 1.0 to 13.0s was used as the range for the rapid gait speed for Ghana and South Africa respectively and thus, leaving 136 outliers from Ghana and 119 from South Africa, and  86 outliers from Ghana and 58 outliers from South Africa for the normal and rapid gait speeds respectively. Other minor data cleaning and validation processes such as selection of relevant variables, and finding the average blood pressure and pulse reading from a set of three were carried.

The two data sets were then combined yielding a total of 8501 entries.


### Overview of Demographics and Socioeconomic Indices for Sub Saharan Africa (Ghana and South Africa combined)

The majority of respondents (~84% representing 8564 respondents) were aged 50 years and above, and 1224 were from the control group (ages 18 to 49).  This represents ~ 1:7 ratio of control group to study group.

The findings below are for the study group only.

Median age was 62 for both Ghana and 61 for South Africa. The lowest count for the age groupings was for 80 years and above. The majority of respondents were female. 
The majority of respondents in Ghana resided in rural areas whereas for South Africa, the opposite was noted. The majority of respondents were either cohabiting or "currently married" in both countries, with the next highest proportion widowed. About twice the proportion of respondents who were separated/divorced in South Africa were separated/divorced in Ghana, and about nine times the proportion of respondents who had never married in Ghana had never married in South Africa. Income was almost equally distributed across the 5 quintiles.

```{r, a}

```

### Overview of Health-related Indices for Sub Saharan Africa (Ghana and South Africa combined)  

<!-- The majority of respondents in Ghana were of normal weight whereas the majority from South Africa were obese.The majority of respondents had no history of chronic disease or alcohol use. For tobacco use however, the majority from Ghana had a positive history whereas the majority from South Africa had a negative history. -->

```{r, com}

```

### In-country Differences

#### Ghana Data

Focusing on Ghana, the median age was 64 for females and 61 for males. The majority of males were "currently married" whereas the majority of females were widowed. Males had a slightly higher normal gait speed than females. With rapid gait speed, the gender difference increased to 0.2. The majority of males had a history of tobacco use whereas majority of females had no history.

#### South Africa Data

Focusing on South Africa, the median age was 60 for females and 59 for males. The majority of males were "currently married" whereas the majority of females were widowed. Males had a slightly higher normal gait speed than females. With rapid gait speed, the gender difference increased slightly further.

```{r, b}

```


```{r c}

```

### Comparison of Gait Speed Across Demographic Variables

The table below shows the multiple regression results with the gait speed as the outcome variable and the various demographics variables as the independent variables.

```{r d}

```


### Comparison of Gait Speed Across Health-Related variables

<!-- Obesity,defined as a Body-Mass Index (BMI) of 30kg/m2 and above, decrease in mean gait speed which was considered statistically significant. Being underweight (BMI <18.5kg/m2) was associated with a reduction in mean gait speed by 0.03 (statistically insignificant).   -->

<!-- Generally, having no history of chronic disease was associated with higher gait speed, with the exception of Diabetes Mellitus, Injury from Road Traffic Accidents and Cataracts. A history of alcohol use was associated with a statistically significant lower gait speed. On the other hand, a history of tobacco use was associated with an insignificant increase in gait speed. -->


```{r, e}

```

#### Association of health indices with gait speed  

Using the 12-item World Health Organization’s Disability Assessment Schedule, version 2 (WHODAS-II), respondent’s functional ability was measured. The 12-item questionnaire measured the activities of daily living(ADLs) and instrumental activities of daily living (IADLs).

The WHODAS-related variables are a series of questions that start with the following preamble: “Overall in the last 30 days, how much difficulty…”, followed by various parameters/tasks.
The respondents' answers were re-coded with **0 equivalent to "none", 1 equivalent to "mild", 2 to "moderate", 3 to "severe" and 4 to "extreme".** Across the 12 variables, a respondent could score a **minimum of 0 and a maximum of 48.** The table below compares the 12-item WHODAS variables in general, and specifically the three questions which deal with the ADLs, as well as height, and BMI and gait speed.

```{r, n}
 
```

### Trend of Gait Speed With Age
It can be seen clearly in the graphs below that there is a decline in gait speed with increasing age:
```{r, g}
 
```

### Zooming In To 75+
The black line represents the linear regression model with the shaded area representing the 95% confidence interval. 
```{r, h}

```


### Gender vs Normal and Rapid Gait Speed
The results below suggest that the male gender is associated with slightly higher mean gait speed:
```{r, i}

```


### Respondents' subjective assessment of pain, health and energy levels

### Association of Health Rating with gait speed

The graph below shows respondents' subjective assessment of health level and it's association with gait speed. The line in the box represent the median, and the entire box represents the interquartile range. The density of the surrounding dotted area represents the number of respondents' who chose that answer as their health rating.

```{r, ps}

```

```{r, pe}

``` 