---
title: "SAGE Gait Study"
author: '*Boadiwaa*'
date: "Created on `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(knitr)
library(tidyverse)
library(foreign)
library(ggpubr)
library(gtsummary)
library(modelsummary)
knitr::read_chunk('edits_gait.R')

```

#### Overview of Demographics, Socioeconomic and Health Indices for Sub Saharan Africa (Ghana and South Africa combined)

```{r, a}

```


#### Ghana Data
```{r, b}

```

#### South Africa Data
```{r c}

```

#### Comparison of Mean Gait Speed Across Demographic Variables
```{r d}

```


#### Comparison of Mean Gait Speed Across Health-Related variables
```{r, e}

```


#### Effect of WHODAS variables on gait speed
```{r, m}

```

```{r, n}

```



#### Combined Linear Regressions Table
```{r, f}

```


### Graphs

#### Trend of Gait Speed With Age
It can be seen clearly that with increasing age, the mean gait speed decreases:
```{r, g}

```

#### Zooming In To 60+
```{r, h}

```


#### Gender vs Normal and Rapid Gait Speed
The results below suggest that the male gender is associated with higher gait speed:
```{r, i}

```

```{r, j}

```

```{r, k}

```

```{r, l}

```

